---
title: "Visualization Principles"
author: "Yihun Zeleke"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse) # Collection of tidyverse packages
library(dslabs) # From Edex and Harvard Data Science Program
library(forcats) # Working on Categorical variables

```

### Good to Know

 - Visual cues for encoding data include position, length, area, brightness and color hue.
 - Position and length are the preferred way to display quantities, followed by angles, which are preferred over area. Brightness and color are even harder to quantify but can sometimes be useful.
 - Pie charts represent visual cues as both angles and area, while donut charts use only area. Humans are not good at visually quantifying angles and are even worse at quantifying area. Therefore pie chart and donut charts should be avoided-use a bar plot instead. If you must make a pie chart, include percentage as labels.
 - Bar plots represent visual cues as position and length. Humans are good at visually quantifying linear measures, making bar plots a strong alternative to pie or donut charts.
 
#### Know When to Include *Zero*

- When using bar plots, always start at `0`. It is deceptive not to start at `0` because bar plots imply length is proportional to the quantity displayed. Cutting off the `y-axis` can make differences look bigger than they actually are.
- When using position rather than length, it is not necessary to include `0`(`Scatter plot`, `dot plot`, `boxplot`)

#### Do Not Distort Quantities  

- Make sure your visualizations encode the correct quantities. 
- For example, if you are using a plot that relies on circle area, make sure the area(rather than the radius) is proportional to the quantity. 

#### Order by a Meaningful value 

- It is easiest to visually extract information from a plot when ordered by a meaningful value. The exact value on which to order will depend on your data and the message you wish to convey with your plot.
- The default ordering for categories is alphabetical if the categories are strings or by factor level if factors. However, we rarely want alphabetical order.

#### Show the Data 

- A dynamic plot a bar graph of group averages with error bars denoting standard errors-provides almost no information about a distribution.  
- By showing the data, you provide viewers extra information about distributions.  
- Jitter is adding a small random shift to each point in order to minimize the number of overlapping points. To add jitter, use the `geom_jitter()` geometry instead of `geom_point()`.  
- Alpha blending is making points somewhat transparent, helping visualize the density of overlapping points. Add an `alpha` argument to the geometry.

#### Use Common Axes 

- Ease comparison by keeping axes the same when comparing data across multiple plots. 
- Align plots vertically to see horizontal changes.Align plots horizontally to see vertical changes.
- Bar plots are useful for showing one number but not useful for showing distributions.

#### Consider Transformations

- Use transformations when warranted to ease visual interpretation. 
- The log transformation is useful for data multiplicative changes. The logistic transformation is used for fold changes in odds. The square root transformation is useful for count data.

#### Compared Visual Cues Should Be Adjacent 

- When two groups are to be compared, it is optimal to place them adjacent in the plot.
- Use color to encode groups to be compared. 
- Consider using a color blind friendly palette like the one in this video.

```{r}
color_blind_friendly_cols <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

p1 <- data.frame(x = 1:8, y = 1:8, col = as.character(1:8)) %>%
    ggplot(aes(x, y, color = col)) +
    geom_point(size = 5)
p1 + scale_color_manual(values = color_blind_friendly_cols)
```

#### Slope Charts

- Consider using a slope chart or Bland-Altman plot when comparing one variable at two different time points, especially for a small number of observations. 
- Slope charts use angle to encode change. Use geom_line() to create slope charts. It is useful when comparing a small number of observations. 
- The Bland-Altman plot (Tukey mean difference, MA plot) graphs the difference between conditions on the y-axis and the mean between conditions on the x-axis. It is more appropriate for large numbers f observations than slope charts. Also shows the difference versus the average.


```{r}
west <- c("Western Europe", "Northern Europe", "Southern Europe", "Northern America", "Australia and New Zealand")
data <- gapminder %>% 
    filter(year %in%c(2010, 2015) & region %in%west & !is.na(life_expectancy) & population > 10^7) 


data %>% 
    mutate(location = ifelse(year == 2010, 1, 2),
           location = ifelse(year == 2015 & country%in%c("United Kingdom", "Portugal"),
                             location + 0.22, location),
           hjust = ifelse(year == 2010, 1,0)) %>% 
    mutate(year = as.factor(year)) %>% 
    ggplot(aes(year, life_expectancy, group = country))+
    geom_line(aes(color = country), show.legend = FALSE)+
    geom_text(aes(x = location, label = country, hjust = hjust),
              show.legend = FALSE)+
    xlab("")+
    ylab("Life Expectancy")
```

<br>
**Bland-Altman plot** 
```{r}
library(ggrepel)
data %>% 
    mutate(year = paste0("life_expectancy_",year)) %>% 
    select(country, year, life_expectancy) %>% 
    spread(year, life_expectancy) %>% 
    mutate(average = (life_expectancy_2015 + life_expectancy_2010)/2,
           difference = life_expectancy_2015 - life_expectancy_2010) %>% 
    ggplot(aes(average, difference, label = country))+
    geom_point()+
    geom_text_repel()+
    geom_abline(lty = 2)+
    xlab("Average of 2010 and 2015")+
    ylab("Difference between 2015 and 2010")
```
<br>
#### Encoding a Third Variable 

- Encode a categorical third variable on a scatterplot using color hue or shape. Use the shape argument to control shape.
- Encode a continuous third variable on a using color intensity or size. 

#### Case Study: Vaccines 

- Vaccines save millions of lives, but misinformation has led some to question the safety of vaccines. The data support vaccines as safe and effective. We visualize the data about measles incidence in order to demonstrate the impact of vaccination programs on disease rate.
- The `RcolorBrewer` package offers several color palettes. Sequential color palettes are best suited for data that span from high to low. Diverging color palettes are best suited for data that are centered and diverge towards high or low values. 
- The geom_tile() geometry creates a grid of color tiles.
- Position and length  are strong cues than color for numeric values, but color can be appropriate sometime.

```{r}
data("us_contagious_diseases")
str(us_contagious_diseases)
```

```{r}

the_disease = "Measles"
mel_dat <- us_contagious_diseases %>% 
    filter(!state%in% c("Alaska","Hawaii") & disease == the_disease) %>% 
    mutate(rate = count/population*10000) %>% # The Rate is per 10,000 people
   mutate(state = fct_reorder(state, rate))
```


```{r}
mel_dat %>% 
    filter(state == "California" & !is.na(rate)) %>% 
    ggplot(aes(year , rate))+
    geom_line()+
    geom_vline(xintercept = 1963, color = "blue")+
    ylab("Cases per 10,000")+
    ggtitle("Vaccine was started in California")
```
<br>
*Can we show data for all states in one plot? By using year, rate, and state?*  
We choosing colors to quantify a numeric variable, we have to choose two options of color: `sequential` or `divergent`? Sequential palettes are suited for data that goes from high to low. High values are clearly distinguished from the low values. And on the other hand diverging colors are used to represent values that verge from a center. An example of when we would use the divergent pattern to show `heights` and `standard deviations` away.

```{r}

mel_dat %>% 
    ggplot(aes(year, state, fill = rate))+
    geom_tile(color = "grey50")+
    scale_x_continuous(expand = c(0,0))+
    scale_fill_gradientn(colors =  RColorBrewer::brewer.pal(9, "Reds"), trans = "sqrt")+
    geom_vline(xintercept = 1963, color = "blue")+
    theme_minimal()+
    theme(panel.grid = element_blank())+
    ggtitle(the_disease)+
    ylab("")+
    xlab("")

```
<br>
However, one limitation of this plot is that is uses color to represent quantity, which we earlier explained makes it a bit harder to know exactly how high it is going.  

Position and length are better cues. If we are willing to lose data information,we can make a version of the above plot that shows the values with position. We can also show the average for the US.

```{r}

avg <- us_contagious_diseases %>% 
    filter(disease == the_disease) %>% 
    group_by(year) %>% 
    summarise(us_rate = sum(count, na.rm = T)/sum(population, na.rm = T)*10000)

mel_dat %>%
    filter(!is.na(rate)) %>%
    ggplot() +
    geom_line(aes(year, rate, group = state), color = "grey50", 
        show.legend = FALSE, alpha = 0.2, size = 1) +
    geom_line(mapping = aes(year, us_rate), data = avg, size = 1, col = "black") +
    scale_y_continuous(trans = "sqrt", breaks = c(5, 25, 125, 300)) +
    ggtitle("Cases per 10,000 by state") +
    xlab("") +
    ylab("") +
    geom_text(data = data.frame(x = 1955, y = 50),
        mapping = aes(x, y, label = "US average"), color = "black") +
    geom_vline(xintercept = 1963, col = "blue")
```
<br>

#### Avoid Pseudo and Gratuitous 3D Plots 

- In general, pseudo-3D plots and gratuitous 3D plots only add confusion. Use regular 2D plots instead.

#### Avoid Too Many Significant Digits  

- In tables, avoid using too many significant digits. Too many digits can distract the meaning of your data. 
- Reduce the number of significant digits globally by setting ab option. For example, `options(digits = 3)` will cause all future computations that session to have 3 significant digits.
- Reduce the number of digits locally using `round() `or `singif()`.































